\section{Design}
\label{sec:design}


\input{figures/sec-3-workflow}
%---------------------------------------------------------------------------------------------------------
A verifier rejection marks an operation at which the verifier cannot establish a required proof. Its terminal error identifies this \emph{rejection location}, but not necessarily where the proof was lost or where the source should be changed. From an existing verifier log, \sys reconstructs the observable lifecycle of the required proof. We distinguish the rejection location, an observed \emph{proof-break witness} after which logged evidence no longer supports the proof, and the developer-chosen \emph{repair location}. A witness narrows the diagnosis, but is neither a minimal causal instruction set nor necessarily the repair location.

Figure~\ref{fig:design-overview} summarizes the three-stage workflow: log analysis normalizes evidence around the rejection, proof reconstruction identifies and tracks the required proof, and diagnostic generation turns the resulting evidence into guidance the developer can act on.

\subsection{Architecture and Data Flow}
\label{sec:design-architecture}

\sys is a log-first pipeline whose only required input is an existing verifier log. Optional source or object metadata improves instruction-to-source correlation but is not needed for log-level diagnosis.

The Log Analysis stage isolates the final rejection and extracts the terminal error, the per-instruction abstract state and branch information, instruction identifiers, and available source annotations into a normalized evidence stream. The Proof Reconstruction stage maps the terminal error to a proof family---such as pointer provenance, packet bounds, scalar range, non-nullness, stack initialization, or reference lifetime---instantiates the required proof at the rejected operation, and tracks evidence for that proof. Its evidence record contains the required proof, rejection location, supporting evidence, observed establishment or loss events, and a proof-break witness when the log exposes one. The Diagnostic Generation stage selects primary and related spans, records their localization precision, derives a lifecycle-specific next action, and may attach a failure class and confidence when supported by ordered evidence rules.

The result is a Rust-style plain-text diagnostic with a stable error identifier, the required proof, relevant spans and evidence, a witness when observable, and \texttt{help:} guidance.

\subsection{Reconstructing the Verifier-Visible Proof}
\label{sec:design-proof}

\sys first identifies which proof the verifier required. The terminal error anchors the rejected operation and immediate symptom, while nearby abstract state supplies parameters such as the affected register, access offset and size, scalar range, pointer type, or reference identifier. Together they instantiate the required proof at the rejection location, without assuming the rejection location is the eventual repair location. For example, in Figure~\ref{fig:overview}, \texttt{R5 invalid mem access `scalar'} maps to a pointer-provenance proof. From the rejected load and nearby state, \sys infers that \texttt{R5} must still carry verifier-recognized packet-pointer provenance when the rejected load reads it.

Proof-family-specific extraction records three event types: a \emph{proof-establishment event}, where logged evidence explicitly supports the proof; a \emph{proof-loss event}, where a later logged state is incompatible with it; and the \emph{rejection event}, where the verifier requires but cannot establish it. A missing snapshot or unprinted state is not itself proof loss. Each event retains its program location, relevant state, evidence source, and recoverable source span. When both a supporting and a later incompatible observation exist, their transition becomes the proof-break witness; otherwise, \sys reports only the required proof and rejection location.

In the running example, an earlier state represents \texttt{R5} as \texttt{pkt(off=34,r=42)}, whereas a later state before instruction~37 represents it as scalar value \texttt{40}. The final load dereferences \texttt{R5}, so \sys reports the observed pointer-to-scalar transition as the witness and instruction~37 as the rejection location. This witness marks only where provenance ceased to be visible in the printed evidence; it does not identify a unique causal instruction, account for unprinted paths, or determine the repair location.

\subsection{Evidence-Aware Diagnostic Synthesis}
\label{sec:design-synthesis}

Diagnostic synthesis adapts its guidance to the reconstructed proof lifecycle. \sys localizes the evidence needed for repair: the rejected operation, the required proof at that operation, the proof-break witness when the log exposes one, and the span and next action that guide the developer in re-establishing the proof. If no establishment event is observed, \sys recommends introducing or re-deriving the missing proof, for example through a same-path bounds check, null check, initialization, or reference acquisition. If the proof is visible and later lost, it recommends preserving or re-deriving it near the rejected use. If intermediate evidence is insufficient, it recommends conservative local re-establishment without claiming an unsupported witness.


The evidence record also determines localization. \sys selects the most informative location as the primary span, reports relevant rejection, establishment, or loss locations as related spans, and states whether each is backed by source mapping, a BPF instruction identifier, or only a log position.

\sys may also attach a failure class and confidence when ordered evidence rules support one. These labels summarize the diagnostic but do not drive proof reconstruction or span selection. In particular, an observed proof transition alone does not determine whether the repair belongs in source, compiler lowering, loader relocation, or the verifier. When the evidence is insufficient, \sys keeps the evidence-backed diagnostic without forcing an attribution.

Developers use the resulting guidance to choose and validate a repair through the normal compile-and-load workflow. \sys neither rewrites the program nor treats the witness as the repair location.