\section{Evaluation}
\label{sec:evaluation}

This section evaluates the \sys diagnostic with two case studies and a repair measurement.
% 本节用两个 case study 和一个修复测量来评估 \sys 的 diagnostic。
We first run two case studies on \empiricalset: one rejection where the diagnostic recovers the repair information the terminal error omits, and a pair where it separates the responsible layer behind a shared terminal error.
% 我们先在 \empiricalset 上做两个 case study：一个 rejection 展示 diagnostic 从 terminal message 里恢复出被省略的修复信息，一对 rejection 展示它如何区分一条共享 message 背后的责任 layer。
We then measure on \bench whether replacing the raw verifier log with the \sys diagnostic raises downstream repair, counting only the fixes that pass tests independent of \sys.
% 然后我们在 \bench 上测量：把 raw verifier log 换成 \sys diagnostic 是否提升下游修复，只数那些通过独立于 \sys 的测试的 fix。

\noindent\textbf{Setup.} The two case studies draw on \empiricalset, the 235 reproduced rejections of \S\ref{sec:study:corpus}.
% 两个 case study 用的是 \empiricalset，也就是 \S\ref{sec:study:corpus} 里那 235 个复现 rejection。
The repair measurement uses \bench, 75 source-level repair tasks, 40 built around a required verifier proof and 35 minimized from open-source projects such as Cilium~\cite{cilium}, xdp-tools~\cite{xdptools}, and bpftime~\cite{zheng2025extending}.
% 修复测量用 \bench：75 个源码级修复任务，其中 40 个围绕 verifier proof obligation 构造、35 个从 Cilium、xdp-tools、bpftime 这类开源项目最小化而来。
Each task ships the buggy program with an executable test suite independent of \sys, which accepts a fix only if it loads through the kernel verifier and passes the case's functional and source-semantics checks.
% 每个任务都带出错程序和一个独立于 \sys 的可执行测试套件，它只在修复通过内核 verifier 加载、并通过该 case 的功能和源码语义检查时才接受。

\subsection{Recovering the Missing Repair Information}
\label{sec:eval-diagnostic-completeness}

The verifier's terminal error names the operation where verification stopped, but a repair needs more: which proof the program failed to keep, and where it lost that proof.
% verifier 的 terminal message 点出 verification 停下的那个 operation，但修复需要更多：程序没能保住哪条 proof、又在哪丢的。
We examine one rejection from the aya project.
% 我们细看 aya 项目里的一个 rejection。

As shown in Figure~\ref{fig:diag-example}, the program treats the map object \texttt{\&globals} as a map-value pointer, casting its address to \texttt{\_\_u64 *} and reading and writing through the result.
% 如图~\ref{fig:diag-example} 所示，程序把 map 对象 \texttt{\&globals} 当成 map-value 指针，把它的地址 cast 成 \texttt{\_\_u64 *}，再通过结果读写。
The verifier rejects the write with the terminal line \texttt{only read from bpf\_array is supported}.
% verifier 用末行 \texttt{only read from bpf\_array is supported} 拒绝这次写。
The line is correct, but it gives only the symptom: the verifier sees a write through the map object where a map value is required.
% 这行没错，但只给了 symptom：verifier 看到的是通过 map 对象本身的写，而那里需要的是一个 map value。

From the same log, \sys reconstructs the lost proof.
% 从同一份 log，\sys 重建出丢失的 proof。
The diagnostic states the proof the rejected write needs, a map-value pointer derived through a map helper before any access to map contents.
% diagnostic 点明被拒的写需要的 proof：在读写 map 内容前，先用 map helper 派生出一个 map-value 指针。
It then localizes the loss to the cast that produced a raw map-object pointer.
% 然后它把丢失点定位到那个产生原始 map-object 指针的 cast。
As shown in Figure~\ref{fig:diag-example}, the developer's fix follows that proof: it calls \texttt{bpf\_map\_lookup\_elem}, checks the returned pointer, and updates the value through the result.
% 如图~\ref{fig:diag-example} 所示，开发者的修复正照这条 proof：调用 \texttt{bpf\_map\_lookup\_elem}、检查返回指针、再通过结果更新值。
The terminal line gives only the rejected operation, while the diagnostic recovers the lost proof and the point where the program lost it.
% terminal line 只给出被拒的 operation，而 diagnostic 恢复了丢失的 proof、以及程序丢失它的那个位置。

\input{figures/sec-5-diagnostic-example}

\subsection{Separating the Responsible Layer}
\label{sec:eval-layer-attribution}

A second diagnostic gap is that one terminal verifier message can call for repairs in different layers.
% 第二个 diagnostic gap 是：同一条 terminal verifier message 可能对应不同 layer 的修复。
\sys reports the responsible layer as a best-effort attribution, labeling a rejection \texttt{source\_bug} for the source layer of \S\ref{sec:study} or \texttt{lowering\_artifact} for the compiler layer. We inspect whether \sys separates two rejections that share one terminal error but belong to different layers.
% \sys 把责任 layer 作为 best-effort 归因报告，所以我们检视它能否区分两个共享同一条 message、却属于不同 layer 的 rejection。

Figure~\ref{fig:layer-attr} shows two rejections that both end with \texttt{invalid mem access `scalar'}.
% 图~\ref{fig:layer-attr} 给出两个都以 \texttt{invalid mem access 'scalar'} 结尾的 rejection。
In Figure~\ref{fig:layer-attr}(a), the program constructs a packet pointer by casting an integer offset, \texttt{(void *)(sizeof(*eth) + nh\_off)}.
% 在图~\ref{fig:layer-attr}(a) 中，程序用一个整数 offset 强转出 packet 指针：\texttt{(void *)(sizeof(*eth) + nh\_off)}。
The resulting value has no verifier-tracked packet base, so the later packet-field load uses a scalar where the verifier requires a real pointer.
% 得到的值没有 verifier 跟踪的 packet base，于是后面那次 packet 字段 load 在需要真指针的地方用了一个 scalar。
\sys reports \texttt{source\_bug}: the source never establishes the pointer provenance the rejected load requires.
% \sys 报 \texttt{source\_bug}：源码从未建立被拒 load 所需的指针 provenance。

In Figure~\ref{fig:layer-attr}(b), the source expresses the intended packet-bound proof, deriving \texttt{udph} from an IPv4 or IPv6 header and bounds-checking it against \texttt{data\_end} before the read.
% 在图~\ref{fig:layer-attr}(b) 中，源码表达了它想要的 packet-bound proof：从 IPv4 或 IPv6 头派生出 \texttt{udph}，在读取前对着 \texttt{data\_end} 做边界检查。
The rejection arises after compiler lowering merges the branch and hides the verifier-recognized packet-pointer type before the load.
% 这次 rejection 是在编译器 lowering 合并分支、在 load 前把 verifier 认得的 packet 指针类型藏掉之后才出现的。
The upstream fix steers the compiler and leaves the source logic unchanged, so the layer label holds independently of \sys.
% 上游的修复是去引导编译器、不动源码逻辑，所以这个 layer 标签独立于 \sys 也成立。
\sys reports the case as \texttt{lowering\_artifact}.
% \sys 把这个 case 报为 \texttt{lowering\_artifact}。
On these two rejections, \sys reports a different layer for each, separating a source bug from a lowering artifact behind one terminal error.
% 在这两个 rejection 上，\sys 各报了不同的 layer，把一条 terminal message 背后的 source bug 和 lowering artifact 区分开。

\input{figures/sec-5-layer-attr}

\subsection{Repair Effectiveness at Scale}
\label{sec:eval-repair-effectiveness}

The two case studies show that \sys exposes the repair information the terminal verifier message omits.
% 两个 case study 表明 \sys 暴露了 terminal verifier message 省略掉的修复信息。
We now test whether that information improves end-to-end repair on \bench.
% 现在我们测试这些信息是否提升 \bench 上的端到端修复。
For each task, a model receives the buggy program with either the raw verifier log or the \sys diagnostic and writes a candidate fix for the test suite to judge.
% 对每个任务，模型拿到出错程序加 raw verifier log 或 \sys diagnostic 之一，写出一个候选 fix 交给测试套件判定。
Because the tests are independent of \sys, a fix earns credit only by passing the same checks a correct program must pass.
% 因为测试独立于 \sys，一个 fix 只有通过一个正确程序同样要过的那些检查才算数。

We score each task in two modes.
% 我们用两种模式给每个任务打分。
The one-shot mode tests the diagnostic on its own, and the retry mode allows one failure-informed retry to test whether the help persists.
% one-shot 模式测 diagnostic 单独是否有用；retry 模式允许一次带失败反馈的重试，测帮助是否还在。
Our primary model is Qwen3.6 27B, and we also run the hosted GLM 5.2 and the smaller Qwen2.5 3B at temperature zero.
% 主模型是 Qwen3.6 27B，另外在 temperature 0 下跑托管的 GLM 5.2 和更小的 Qwen2.5 3B。
We include the small 3B model to see whether the gain survives at low model capacity.
% 我们加入小的 3B 模型，看这个提升在低模型容量下是否还在。

Replacing the raw verifier log with the \sys diagnostic raises repair for every model, as shown in Figure~\ref{fig:bpfix-bench}.
% 把 raw verifier log 换成 \sys diagnostic，每个模型的修复都上升，如图~\ref{fig:bpfix-bench} 所示。
On Qwen3.6 27B, one-shot repair rises from 22 of 75 cases with the raw log to 38 with the diagnostic, and one retry raises the two counts to 30 and 44.
% 在 Qwen3.6 27B 上，one-shot 修复从 raw log 的 75 例中 22 升到 diagnostic 的 38，一次重试把两者抬到 30 和 44。
GLM 5.2 shows the same pattern, from 28 to 38 in one shot and 52 against 47 with the retry.
% GLM 5.2 走向一致：one-shot 从 28 到 38，retry 时 52 对 47。
The small Qwen2.5 3B repairs none of the 75 from the raw log, and 8 from the diagnostic in one shot and 10 with one retry.
% 小模型 Qwen2.5 3B 从 raw log 一个都修不出，从 diagnostic 在 one-shot 修出 8 个、retry 10 个。
The gain holds across all three models, which points to the diagnostic as its source.
% 这个提升在三个模型上都成立，说明来源是 diagnostic。

\input{figures/sec-5-bpfix-bench}

To locate the gain, we charge each failed one-shot candidate to the first stage it fails: returning a program, compiling, loading through the verifier, passing the functional test, or passing the source-semantics test.
% 为定位提升来自哪里，我们把每个失败的 one-shot 候选归到它第一个失败的阶段：返回程序、编译、过 verifier 加载、过功能测试、过源码语义测试。
As Table~\ref{tab:failure-stage} shows, the diagnostic mainly reduces verifier-load and source-semantics failures.
% 如表~\ref{tab:failure-stage} 所示，diagnostic 主要减少 verifier 加载失败和源码语义失败。
On Qwen3.6 27B, load failures fall from 19 to 10 and source-semantics failures from 22 to 16, and GLM 5.2 shows the same two reductions, from 10 to 5 and from 25 to 22, with compile failures staying low for both models.
% 在 Qwen3.6 27B 上，加载失败从 19 降到 10、源码语义失败从 22 降到 16；GLM 5.2 是同样两处下降，从 10 到 5、从 25 到 22，两个模型的编译失败都很低。

For the small Qwen2.5 3B, the raw log is often too weak a repair signal: 62 one-shot candidates still fail at verifier load, and three raw-log prompts exceed the model's context window and return no program.
% 对小模型 Qwen2.5 3B，raw log 常常是太弱的修复信号：62 个 one-shot 候选仍在 verifier 加载处失败，还有三个 raw-log prompt 超出模型上下文窗口、没返回程序。
With the shorter \sys diagnostic, model-call failures disappear and verifier-load failures fall to 39, though compile failures rise from 7 to 14.
% 换成更短的 \sys diagnostic 后，模型调用失败消失、verifier 加载失败降到 39，不过编译失败从 7 升到 14。
\sys gives even a small model enough verifier-facing information to attempt more meaningful repairs, while ordinary code-generation errors remain.
% \sys 让即便小模型也拿到足够的 verifier 向信息去尝试更有意义的修复，而普通代码生成错误依然存在。

\input{tables/sec-5-failure-stage}

Overall, the quantitative results match the two case studies.
% 总体上，定量结果和两个 case study 一致。
The diagnostic helps most at the stages where a repair must restore the verifier-visible proof, loading through the verifier and preserving the program's source semantics.
% diagnostic 在"修复必须恢复 verifier 可见事实"的阶段帮助最大：过 verifier 加载、以及保住程序的源码语义。
These stages are exactly where \sys localizes the lost proof, so the model receives what a repair must restore.
% 这些阶段正是 \sys 定位丢失 proof 的地方，所以模型拿到的正是修复需要恢复的东西。
